### 2023-01-27 doc/manuscript/translocation.qmd
Minor edits to introduction outline

Severe effects of disease on global biodiversity (bats, primates, amphibians, corals, sea stars, etc.)

-   Species recovery in the face of severe disease (population growth rate \< 1 ) will depend on the ability of a species to move or change.

@@

We estimated survival of translocated frogs, recruitment of new frogs into the adult population, and population size using a Bayesian open-population Jolly-Seber mark-recapture model with known additions (i.e., translocated cohorts) to the population, as implemented by the mrmr package [@joseph2019] and using R Statistical Software \[v4.4.4; \@rsoftware2022\]. The model tracks the states of *M* individuals that comprise a superpopulation made up of real and pseudo-individuals [see @joseph2018 for details]. The possible states of individuals include "not recruited", "alive", and "dead". The possible observations of individuals include "detected" and "not detected". We assume that individuals that are in the "not recruited" or "dead" states are never detected (i.e., there are no mistakes in the individual PIT tag records). For all models, we used mrmr defaults for priors, number of chains (4), and warmup and post-warmup iterations (2000 for each). We evaluated convergence of the Markov chain Monte Carlo (MCMC) algorithm using trace plots and Gelman-Rubin statistics (Rhat). Code used to construct CMR datasets and fit models for each translocated population is available from a GitHub repository ([https://github.com/SNARL1/cmr-analysis](https://github.com/SNARL1/cmr-analysis.)).

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:--------------------------------------------------------------------------|:---------------------|:-------------------------------------------------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |

>>>>

P1: Emerging infectious diseases pose a severe threat to human, livestock, and wildlife populations.

-   In wildlife populations, disease has been responsible for dramatic declines, extirpations, and even extinctions in mammals, birds, echinoderms, amphibians, and other taxa (Scheele, avian malaria, black-footed ferrets, sea stars, etc.).

-   While the dramatic and immediate nature of disease-induced declines in wildlife has led to intensive focus on documenting, understanding, and mitigating disease impacts (citations), populations previously decimated by disease are also beginning to recover (citations).

-   However, given the nascent nature of wildlife recoveries in the presence of disease, we know far less about the mechanisms driving population recovery and how these mechanisms can be leveraged to re-establish host populations across large spatial scales.

-   In particular, for species of conservation concern as a result of disease-induced declines, demonstrating the long-term viability and feasibility of potential recovery actions in the presence of disease and elucidating the mechanisms underlying long-term recovery success remain largely absent from conservation biology (but see XX).

P2: Host evolution following disease-induced declines is one mechanisms through which hosts can recover (Carlson et al. 2014, TREE, Searle et al. 2020).

-   Evolutionary rescue describes natural host evolution whereby rapid evolutionary changes originating from variation within a host population reverse ongoing population declines (Carlson).

-   In host-pathogen systems, evolutionary rescue is more likely to occur when hosts experience repeated selection events (e.g., through re-infections), there is a persistent pathogen reservoir or frequency-dependent infection dynamics, and populations experience increased per capita birth rates at small population sizes (Golas et al. 2021).

-   Important for conservation action, evolutionary rescue not only promotes recoveries of focal populations, but can also provide sources of resistant individuals to reestablish populations extirpated following disease emergence (Mendelson et al. 2019, Joseph and Knapp 2018).

P3: For resistant populations to contribute to landscape-scale species recovery, the fitness benefits gained through selection should manifest beyond the natal environment, allowing for the long-term growth and persistence of incipient populations in novel environments.

-   Translocations of individuals from putatively resistant populations to formerly-occupied-but-now-extirpated habitats provides a way to disentangle the contributions of natal environment and host genetics on host persistence and recovery.

-   Something here about previous work with translocations and disease.

-   Simply put, resistant hosts that demonstrate disease resistance, population growth, and long-term viability in non-natal environments provide a means to recover disease-impacted species at a landscape-scale.

P4: Description of MYL frog species complex, impacts of Bd, and evidence of recent recovery.

P5: Here, we combine X translocations of resistant MYL populations, X longitudinal surveys, mark-recapture analysis, and population modeling to demonstrate that resistant populations of an endangered amphibian species can recover and persist across a landscape of disease.

-   We then use genetic evidence to demonstrate that observed disease resistance is likely a result of selection and provide five lines of evidence that evolutionary rescue is a strong candidate mechanism for emergence of this resistance.

1.  Epidemiological conditions in MYL-Bd systems X, Y, and Z are conducive to evolutionary rescue.
2.  Differences in allele frequencies exist between adjacent Bd-naive and recovering populations.
3.  There is a shared selective response across recovering populations.
4.  Changes in population growth rate due to evolution extend beyond the natal environment.
5.  In populations established via translocation of frogs from recovering populations, Bd load is not an important predictor of post-translocation frog survival (in contrast to important relationship between Bd load and frog survival during Bd-caused epizootics).

-   As such, we conclude that natural selection combined with translocations promotes the recovery of a species on the brink of disease-induced extinction. These results may have broad implications for the many species worldwide that are threatened with extinction by the emergence of novel diseases.

@@

We estimated survival of translocated frogs, recruitment of new frogs into the adult population, and population size using a Bayesian open-population Jolly-Seber mark-recapture model with known additions (i.e., translocated cohorts) to the population, as implemented by the mrmr package [@joseph2019] and using R Statistical Software [v4.4.4; @rsoftware2022]. The model tracks the states of *M* individuals that comprise a superpopulation made up of real and pseudo-individuals [see @joseph2018 for details]. The possible states of individuals include "not recruited", "alive", and "dead". The possible observations of individuals include "detected" and "not detected". We assume that individuals that are in the "not recruited" or "dead" states are never detected (i.e., there are no mistakes in the individual PIT tag records). For all models, we used mrmr defaults for priors, number of chains (4), and warmup and post-warmup iterations (2000 for each). We evaluated convergence of the Markov chain Monte Carlo (MCMC) algorithm using trace plots and Gelman-Rubin statistics (Rhat). Code used to construct CMR datasets and fit models for each translocated population is available from a GitHub repository ([https://github.com/SNARL1/cmr-analysis](https://github.com/SNARL1/cmr-analysis.)).

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:--------------------------|:------------------|:-------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |
